---
title: "ITEC 621 - Homework 1 - R Practice"
author: "Shawn Newman"
date: "January 20, 2019"
output: html_document
---

```{r global_options, include=FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

1. **Rectangle Area** This function takes two variables as input and prints a sentence using those two variables as the sides of a rectangle.

```{r}
Area = function(x,y){
  Area.result<-x*y
  print(paste("The area of a rectangle with sides ",x," X ",y," is ",Area.result,sep = ""))
}
Area(3,4)
```

2. **Rectangle Loop** This function creates a series of ten sentences giving the area of a rectangle with sides i and i*2 where i = 1 through 10.

```{r}
for(i in 1:10){Area(i,i*2)
}
```

3. **Read .csv data file** This code imports a csv file into RStudio as "Credit" and then prints the first six rows of data followed by the first five rows and first five columns.

```{r}
Credit <- read.table("Credit.csv",header = TRUE,sep = ",")
head(Credit)
Credit[1:5,1:5]
```

4. **lm() model, all variables ** This creates a linear model to explain the credit rating as a function of all available independent variables in "Credit" and then displays the model's summary results.

```{r}
fit.rating<-lm(Rating~Income+Limit+Cards+Age+Education+Gender+Student+Married+Ethnicity+Balance,Credit)
summary(fit.rating)
```

5. **lm() model, 5 variables **

```{r}
fit.rating.5<-lm(Rating~Income+Limit+Cards+Married+Balance,Credit)
summary(fit.rating.5)

```

6. **Object's class** Displays the object class (type of data) for Gender, Income, and Cards

```{r}
class(Credit$Gender)
class(Credit$Income)
class(Credit$Cards)
```

7. **What do these classes mean?** Gender is a factor or categorical variable, this factor has two options in the dataset. Income is a numeric variable that has numbers with values after the decimal place. Cards is an integer value, a variable that only contains whole number values.

```{r}
Object6Classes<-list("Gender"=c("factor",paste("Gender is a factor or categorical variable, this factor has two options in the dataset.")),"Income"=c("integer",paste("Income is a numeric variable that has numbers with values after the decimal place.")),"Cards"=c("numeric",paste("Cards is an integer value, a variable that only contains whole number values.")))
Object6Classes
```

8. **Income.vect** This code isolates the income column and creates a variable with that data called "income.vect," and then displays the first six values of that vector.

```{r}
income.vect<-Credit$Income
head(income.vect)
```

9. **Mean, minimum, maximum, standard deviation and variance** This displays the Mean, Min, max, sd, and var for income.vect.

```{r}
inc.summ<-list("Mean"=mean(income.vect),"Min" = min(income.vect),"Max"=max(income.vect),"SD"=sd(income.vect),"Var"=var(income.vect))
inc.summ
```

10. **Income.stats** This code creates a vector that comtains the Mean, Min, max, sd, and var for Income.vect and then displays this vector.

```{r}
Income.stats<-c(inc.summ$Mean,inc.summ$Min,inc.summ$Max,inc.summ$SD,inc.summ$Var)
Income.stats
```

11. **Vector Names** This code gives names to the values in Income.vect and then displays the vector.

```{r}
names(Income.stats)<-c("Mean","Min","Max","StDev","Var")
Income.stats
```

12. **5% Inflation** This displays the first six values of the income vector after accounting for 5% inflation.

```{r}
income.vect.5<-income.vect*1.05
head(income.vect.5)
```

13. **Income Boxplot** Shows a boxplot for the predictor "Income" within the table Credit.

```{r}
boxplot(Credit$Income)
```

14. **Income by Gender Boxplot** Shows a boxplot of "Income" subsetted by the Gender of the individual. This boxplot shows that the top incomes for females are lower than the top incomes for males.

```{r}
boxplot(Credit$Income~Credit$Gender)
```

15. **Plot Credit Rating vs. Income**

```{r}
plot(Credit$Income,Credit$Rating,xlab = "Income",ylab = "Credit Rating")
```

16. **Credit Rating Histogram** This creates a histogram showing the frequency range of credit ratings.

```{r}
hist(Credit$Rating,main = "Credit Rating Histogram",xlab = "Rating")
```

17. **Credit Rating qqplot** 

```{r}
qqnorm(Credit$Rating)
qqline(Credit$Rating)
```

18. **Is data normally distributed? Why or why not?** This data is not normally distributed, the Credit ratings are primarily in the lower range with a smaller tail in the higher end of the range.

```{r}
```

19. **ggplot2 Histogram** This code creates a histogram of the "Rating" data using ggplot.

```{r}
RatHist<-ggplot(data = Credit, aes(Credit$Rating)) + geom_histogram()
RatHist
```

20. **Dual Line Plot**

```{r}
RatLine<-ggplot(data = Credit, aes(Income,Rating)) + geom_line()
RatLine + facet_wrap(~Gender,ncol=2)
```
